Higher-Order unification for free!

### 2024-05-03 paper/main.tex
rephrase progress beta (+prog-eta-rhs) termination

\newcommand{\progressetaright}{\emph{$\eta$-progress-\rhs}\xspace}
\begin{definition}[\progressetaright]\label{def:progressetaright} A link
\linketaM{\Gamma}{X}{t} is removed from \linkStore when either 1)
$\testmaybeeta~t$ does not hold (anymore) or 2) by $\eta$-contracting $t$ to
$t'$ where $t'$ does not start with the \elpiIn{lam} constructor. In the
first case, $X$ is unified with $t$, and in the second one, $X$ is unified with
$t'$ (under the context $\Gamma$).
\end{definition}
\todo{comment this rule}
\end{comment}

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH{}, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. \progBetaFail guarantees termination since 
  %it makes 
  %\elpiIn{progress}
  it immediately fails and \progressetaright terminates since it performs 
  terminating operations and, if its permises succeed, the considered
  \linketa is removed from \linkStore.\todo{unclear}
%
  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  % \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
\end{proof}

>>>>

\newcommand{\progressetaright}{\emph{$\eta$-progress-\rhs}\xspace}
\begin{definition}[\progressetaright]\label{def:progressetaright} A link
\linketaM{\Gamma}{X}{t} is removed from \linkStore when either 1)
$\testmaybeeta~t$ does not hold (anymore) or 2) by $\eta$-contracting $t$ to
$t'$ where $t'$ does not start with the \elpiIn{lam} constructor. In the
first case, $X$ is unified with $t$, and in the second one, $X$ is unified with
$t'$ (under the context $\Gamma$).
\end{definition}
\end{comment}

@@

\begin{proof}[Proof sketch]
  Let $l = \linkbetaM{\Gamma}{T}{\rhsBeta}$ a \linkbeta in the store \linkStore.
  The progression made by \progBetaRH{} terminates: $l$ is removed from
  \linkStore after having performed terminating instructions. If $l$ is not
  activated by \progBetaRH, then $X$ is a variables. The activation
  of \progBetaLL replaces $l$ with the new \linkbeta 
  $\linkbetaM{\Gamma}{T}{Y_{s_1\dots s_n~s}\appsep t_2\dots t_m}$.
  This progression can be triggered at most $m$ times, since at each time
  the number of terms applied to $X$ decreases. In the end $l_m$
  will be removed from \linkStore after the unification between \lhs with \rhs.
  We also note that each \progBetaLL generates a \linketa, yet, according to
  \cref{lemma:prog-eta-terminates}, this does not affect termination. 
  \progBetaFail terminates since it stop progression with failure.
  Finally, \progressetaright performs 
  terminating operations and, if its permises succeed, the considered
  \linketa is removed from \linkStore, ensuring termination.
  %
  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  % \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
\end{proof}
